Road Segment Prioritization for Bicycle Infrastructure

### 2021-02-02 README.Rmd
transport geog citations

---
title: Road Segment Prioritization for Bicycle Infrastructure
date: "`r Sys.Date()`"
geometry: margin=2cm
bibliography:
  - mybibfile.bib
  - rlbibfile.bib
#linenumbers: true
numbersections: true
abstract: "Understanding the motivators and deterrents to cycling is essential for creating infrastructure that gets more people to adopt cycling as a mode of transport. The aim of this paper is to develop methods to support the prioritisation of cycling infrastructure and cycling network design, accounting for cyclist preferences and the growing emphasis on 'filtered permeability' and 'Low Traffic Neighborhood' interventions internationally. Distance decay, routing, and network analysis techniques are used to examine where future cycling demand is most likely to arise, how such demand could be accommodated within existing street networks, and how to ensure a fair distribution of investment. The city of Manchester is used as a case study, although the methods can be applied to any city, as demonstrated in a reproducible code and open access data accompanying the paper. We find that flexible routing profiles that are adjustable to local needs can avoid road segments that are stressful to cycle on, enabling the identification of networks that balance directness against the need for safe and stress-free routes. We assess two different policy interventions , 'utilitarian' and 'egalitarian' by analyzing the connectivity of the city through community detection. From a policy perspective, our findings indicate that the 'egalitarian' prioritisation model approach is the more adequate to support cost-effective strategic cycle network planning. \\par\\textbf{Keywords:} cycling networks, low-traffic neighborhoods, routing, transport equity"
# keywords: "cycling, low-traffic neighborhoods, potential demand, routing"
output:
  # bookdown::github_document2:
  bookdown::pdf_document2:
  # # bookdown::word_document2:
      toc: false
      extra_dependencies: ["subcaption", "longtable", "float", "multirow", "placeins", "multicol"]
---

@@

<!-- - Intro: possible trim down substantially then add some literature in (a) Calculating Potential Demand [DONE]/ (b) Routing [DONE]/ (c) Road Segment Prioritization -->
<!-- - Community Detection: Where in the document should this section be? It is now before Road Segment Prioritization, but I can merge them -->
<!-- Can go into road segment section. -->
<!-- - Should we use other cities for the sake of comparison (compare communities, road types etc) -->
<!-- - How to add tables properly (they are now added using latex code which only shows up when outputting to pdf) -->
<!-- - Footnotes are sometimes split between pages. How do I keep them on the same page? -->
<!-- - What is the best way to write pseudocode (Steps for Algorithm 1 and 2) -->

@@

> "Inadequate cycling infrastructure discourages cycling and wastes public money. Much cycling infrastructure in this country is inadequate. It reflects a belief, conscious or otherwise, that hardly anyone cycles, that cycling is unimportant and that cycles must take no meaningful space from more important road users, such as motor vehicles and pedestrians - [@departmentgearchange2020]"

@@

A considerable amount of research has been done towards that end. Segregated cycling infrastructure^[*Segregated cycling infrastructure* refers to road space that is allocated to cyclists only, with physical separation to protect cyclists from other modes of transport.] has been shown to increase cycling uptake [@aldred2019impacts; @goodman2014new; @marques2015infrastructure], with the separation from motorized vehicles being key [@winters2011motivators]. Revealed preference of cyclists shows that they are willing to deviate from the most efficient routes in order to commute on safer roads [@crane2017longitudinal]. However, such deviations are only considered if they do not considerably increase route circuitry; behaviour studies have found that the probability of choosing a route decreases in proportion to its length relative to the shortest route [@broach2011bicycle; @winters2010far]. Another defining feature for cycling infrastructure is how well connected it is. Cyclists prefer cohesive infrastructure, particularly when cycling on arterial roads with high levels of motorized traffic [@stinson2003commuter]. The lack of well-connected cycling infrastructure is one of the main obstacles to increasing cycling uptake [@caulfield2012determining]. While direct and cohesive cycling networks have been shown to positively impact cycling rates, density^[making an area's bicycle network denser means adding more cycling routes in the area and thereby giving cyclists more route options] of the cycling network is also vital [@schoner2014missing].

@@

*Optimization* techniques have been used to propose improvements to cycling networks. @mesbah2012bilevel propose a bi-level formulation to optimize allocation of cycling lanes to the network without exceeding a set budget. They account for the effect of cycling lanes on car traffic, and attempt to maximize utilization of said lanes with minimal impact on car travel times. Car usage is not considered by @mauttone2017bicycle, who develop an optimization framework that aims to minimize the total user cost of cycling on the network. The aggregate flow on links^[*link* refers to a road segment throughout this research] is obtained by using shortest paths to route existing cycling demand onto the road network, and the solution is a proposed set of links where cycling infrastructure should be added in order to minimize the overall travel cost of cyclists across the network. The cost of traversing a link is given as a function of its length and whether or not it has cycling infrastructure. The problem has also been solved by attempting to find the minimum cost of improving roadway links to meet a desired level of service (LOS) [@duthie2014optimization]. In this formulation, all origin-destination (OD) pairs need to be connected by roads that meet the desired LOS. A limitation of these problem formulations do not explicitly solve for continuity, which is dealt with using either a constraint specifying that each link with a bike lane should be connected to at least one destination [@mesbah2012bilevel], a constraint on maximum deviation from shortest paths [@duthie2014optimization], or a discontinuity penalty to prioritize connected road segments [@mauttone2017bicycle].

In this paper, continuity is analysed by looking at the connectivity of the network through the graph-theoretic concept of *connected components*. Other academics have taken this approach; @natera2019data study the existing cycling network in terms of its disconnected components and introduce two different algorithms to connect these components by their most critical links and, in doing so, measure the size of the growth of the largest connected component as a function of the kilometers of network added. The concept of connected components is also at the core of the methodology proposed by @olmos2020data. After routing the cycling demand onto the network links, they use percolation theory to filter out the links based on the aggregate flow^[*flow* is used throughout this research to refer to the cycling demand when it is routed onto the road network. The flow on any road segment is the cumulative demand on it, resulting from cyclists commuting between various OD pairs] passing through them, varying the flow threshold for filtering to identify the minimum flow at which the whole city is connected by a giant component. While these approaches deal with continuity better, they look at the network as a whole when attempting to improve it, and in doing so fail to account for equitable distribution of infrastructure.

@@

This research attempts to propose an egalitarian framework for cycling network design. This is done by identifying the different sub-networks that exist within the larger network, and ensuring they all get a fair share of investment. Trip patterns in a city are not uniformly distributed geographically, and *community finding* methods have been used to partition study areas into localized areas that experience a disproportionate number of trips within them. @akbarzadeh2018designing use a modularity maximization approach [@blondel2008fast] on taxi trip data to identify 7 different communities in the city of Isfahan, Iran. An optimization problem is then formulated to connect nodes within each community with cycling infrastructure, with the emphasis being on connectivity within the communities, not between them. @bao2017planning adopt a similar methodology, first identifying communities and then using a greedy network expansion algorithm to simultaneously add links to each community. The link with the highest benefit-cost ratio in each community is selected, and the network is grown by adding neighboring links to the solution until a budget limit is met. The benefit is the flow on the link, and each link is assigned a cost based on current road conditions.

<!-- EDIT! Add more to limitations -->
Our work builds on these *community finding* approaches by proposing a similar greedy network expansion algorithm for cycle network expansion within communities. We incorporate community finding methods for study area partitioning with weighted routing to avoid links that are stressful to cycle on. In doing so, we propose a methodology that accounts for motivators and deterrents to cycling. We propose three sub-methodologies that expand on some of the limitations of previous studies. These limitations include (a) bias inherent when basing network design solely on existing cycling demand, (b) proposing routes that may not correspond to studies on cyclist preference and government policies, and (3) insufficient consideration of the ethical principles underlying the analysis. Section \@ref(calculating-potential-cycling-demand) focuses on calculating potential cycling demand. Section \@ref(routing) focuses on routing the demand onto the road network while accounting for cyclist preferences and government priorities. Section \@ref(community-detection) outlines a methodology for partitioning the study area based on a community finding algorithm and routed cycling demand. Section \@ref(road-segment-prioritization) introduces the network expansion algorithms, and compares an approach grounded in ‘egalitarianism' to one grounded in ‘utilitarianism'.

@@

*Origin-Destination (OD)* data can be obtained from a range of sources, including GPS data, household travels surveys and Census data on work locations.
In areas where observed OD data is unavailable, modelling techniques such as spatial interaction models can be used to estimate travel volumes between zones [@black1995spatial; @martinez2013new; @wilson1971family].
In this paper we use open access data from the UK census [@ONS2011flowdata], which contains aggregate statistics on number of commuters between administrative zones --- Middle layer Super Output Areas (MSOA) --- by mode of travel. MSOAs have an average population of just over 8000 [@ofn2018population].^[
See https://wicid.ukdataservice.ac.uk/ for open access to the OD data.
]
Figures \@ref(fig:potdemhistograms) and \@ref(fig:desirefacetcycling) illustrates the proportion of trips cycled by distance and the geographic extent of the input OD dataset used in this paper.

@@

```{r desirefacetcycling, fig.cap="Current and Potential Cycling Demand", out.width="100%", fig.show='hold', echo=FALSE}
knitr::include_graphics("data/Manchester/Plots/desire_facet_cycling.png")
```

@@

The above points are accounted for by using a weighted road network for routing. This has previously been done by multiplying all road segments without cycling infrastructure by a fixed impedance factor [@mauttone2017bicycle], or by assigning different weights to the road segments proportional to the investment cost of bringing them to an acceptable level of stress for cycling  [@duthie2014optimization].
Perceived stress has been found to be a function of vehicular traffic volume and speed [@sorton1994bicycle], both of which vary predictably with road type. 
@gehrke2020cycling build on this information to use routing impedance factors for road segments that are based on road type and the existence of cycling infrastructure.

@@

<!-- **ADD TABLE - THIS IS BASIC** -->
<!-- ```{r, echo=FALSE, message = FALSE} -->
<!-- weight_profiles <- readxl::read_excel("Paper/paper_tables.xlsx", sheet = "Weighting Profile") -->

<!-- knitr::kable(weight_profiles, -->
<!--              caption = "Weighting Profiles") -->
<!-- ``` -->
\begin{longtable}{lrrr}
\caption{Weighting Profiles}
\label{table:weightprofiles}\\
\multirow{2}{*}{\textbf{OSM Road Type}} & \multicolumn{3}{c}{\textbf{Weighting Profile}}                                                 \\*
                                        & \multicolumn{1}{l}{\textit{Unweighted}} & \multicolumn{1}{l}{\textit{Weighted}} & \multicolumn{1}{l}{\textit{Weighted\_2}}  \endfirsthead
Cycleway                                & 1          & 1                                     & 1                                         \\
Path                                    & 1          & 0.9                                   & 0.9                                       \\
Residential                             & 1          & 0.9                                   & 0.9                                       \\
Service                                 & 1          & 0.9                                   & 0.9                                       \\
Tertiary                                & 1          & 0.9                                   & 0.9                                       \\
Track                                   & 1          & 0.9                                   & 0.9                                       \\
Unclassified                            & 1          & 0.9                                   & 0.9                                       \\
Secondary                               & 1          & 0.8                                   & 0.8                                       \\
Primary                                 & 1          & 0.7                                   & \textcolor{red}{0}                        \\
Trunk                                   & 1          & 0.6                                   & \textcolor{red}{0}                        \\
Motorway                                & 1          & \textcolor{red}{0}                    & \textcolor{red}{0}                       
\end{longtable}

@@

1. **Weighted**: Relatively high impedance on Primary and Trunk roads (to minimize cycling on them).
2. **Weighted\_2**: Avoiding Primary and Trunk Roads completely.

@@

Banning cycling flow completely on trunk and primary roads may result in excessively circuitous paths, as seen in Figure \@ref(fig:boxplotcircuity). When routing using the *weighted* weighting profile in Table \@ref(table:weightprofiles), we see that shortest paths increase by less than 5\% on average from unweighted shortest paths, with the largest increases still below 30\%. When routing on primary and trunk roads is banned (*weighted_2* profile in Table \@ref(table:weightprofiles)), the average increase relative to unweighted shortest paths rises to 10\%, with certain locations experiencing more significant negative effects on accessibility. 
Given that cyclists will only deviate from shortest paths by a certain amount to access better cycling infrastructure (as explained in Section \@ref(introduction)), allowing flow on some stretches of trunk and primary roads is necessary to insure cycling uptake and equitable access to cycling infrastructure. In its new vision for walking and cycling, the Department for Transport acknowledges that minimal segregated stretches of bicycle lanes on main roads will be necessary to avoid circuitous cycling networks [@departmentgearchange2020].

@@

It should be reiterated that the weighting profile used for routing has been developed for the purposes of this study. It creates a hierarchy of road preference that is grounded in cyclist preferences and government plans to create LTNs.
A wide range of weighting profiles could be used, to represent different types of cyclists and road environments, as described in @gehrke2020cycling and @furth2016network.
Sensitivity analysis could be done to determine an optimal weighting profile, but given the variation in city road networks<!-- (Figure \ref{fig:perc_person-km}) -->, this would probably require calibration to the specific city. More accurate routing could be carried out given the availability of road-level data. In such cases we would add additional impedance to specific roads, giving more useful routing results than the current methodology which considers all roads of the same type to be equivalent.

One use-case of such granular data would be to identify roads that serve schools. The Department of Transport notes that the number of school children being driven to school has trebled over the past 40 years [@departmentgearchange2020], and so having cycling infrastructure serving schools is key to achieving the government target of getting more children to cycle. This would not be difficult, as over 75\% of children in the UK live within a 15 minute cycle from their school [@departmentcycleinfradesign2020]. @goodman2019scenarios show that if dutch levels of cycling were achieved in the UK, the \% of children cycling to school could increase from 1.8\% to 41\%. In their typology of cyclists, @dill2013four found that a majority of people who say they would never cycle had never cycled to school, whereas confident cyclists were those most likely to have cycled to school. Getting people to cycle from a young age is therefore key to achieving societal change in commuting habits.

# Road Segment Prioritization

@@

A major challenge facing ‘top-down’ planning approaches is how to incorporate egalitarian principles by fairly distributing investments in cycling infrastructure. 
One way of quantifying this is to split up the city into smaller geospatial areas and target equal investment in each of those areas. This approach could also help ensure that on-the-ground surveys are made by local stakeholders, an important component of the planning process  [@parkin2018designing]. 
Community detection offers us a way to delineate such a split; cyclists are limited in their commuting distance (see Figure \@ref(fig:cyclinghistmanc)), and so trip attractors are more likely to have a local catchment area of cyclists.

@@

<!-- 1. Create links between MSOA centroids and weigh these links by potential cycling demand between them.
2. Use Louvian method to determine optimal number of communities and assign each MSOA centroid to a community.
3. Assign each road link to the same community as the closest MSOA centroid to it. -->
```{r, eval = FALSE}
1. Create links between MSOA centroids and weigh these links by potential cycling demand 
   between them.
2. Use the Louvain method to determine the optimal number of communities and assign each MSOA 
   centroid to a community.
3. Assign each road link to the same community as the closest MSOA centroid to it.
```

@@

```{r communitiesmanchester, fig.cap="Communities Based on Potential Cycling Demand (Manchester)", out.width="75%", fig.show='hold', echo=FALSE}
knitr::include_graphics("data/Manchester/Plots/communities_alternative_Manchester.png")
```

@@

## Algorithm 2: Egalitarian Expansion (Focus on Fair Distribution of Resources)

The first algorithm focuses on connectivity and directness, but not on fairly distributing investment. The latter is not a requirement for increasing cycling uptake, but it is fundamental for spatial equity, as explained in Section \@ref(introduction). This algorithm incorporates the ideal of fair distribution by using community detection to partition the road network.

@@

The results of the community detection are used to evaluate the algorithms. This is done by looking at the *person-km satisfied* as cycling infrastructure is added. Person-km is a measure of the total km cycled on a road segment, so it is the product of the number of potential commuters cycling on that road segment ($flow$) and the length of the segment in km ($l$). For each road segment, the person-km is equal to $flow * l$. In the case of Manchester, Table \@ref(tab:personkmtable) shows that almost half of the person-km is in community 1 , while only 0.5\% of total person-km on the network is in community 5.

@@

Looking at the person-km satisfied (Figure \@ref(fig:growthtotal)), we see that the incremental addition of cycling infrastructure is better distributed between communities using Algorithm 2; equal distribution of investment results in the gain in % of person km satisfied in each community being inversely correlated with the size of the community. In addition, we find that the restrictions imposed by Algorithm 2 on the network expansion do not seem to have a noticeable effect on the city-wide \% of person-km satisfied. Comparing both algorithms, we can see that Algorithm 1 provides only marginally quicker city-wide gains than Algorithm 2.

```{r growthtotal, fig.cap="Comparing Overall (Dashed) and Community Level Person-Km Satisfied (Manchester)", out.width="35%", fig.show='hold', echo=FALSE}
knitr::include_graphics(c(
  "data/Manchester/Plots/Growth_Results/growth_utilitarian_satisfied_km_both_flow_column.png",
  "data/Manchester/Plots/Growth_Results/growth_egalitarian_satisfied_km_both_flow_column.png"
))
```

@@

```{r growth3MapandBar, fig.cap="Results of Alg. 2 (Manchester)", out.width="35%", fig.show='hold', echo=FALSE}
knitr::include_graphics(c(
  "data/Manchester/Plots/Growth_Results/growth_egalitarian_priority_all_FLOW.png",
  "data/Manchester/Plots/Growth_Results/growth_egalitarian_investment_highways_flow.png"
))
```

@@

```{r componentsandGCC, fig.cap="Network Characteristics", out.width="35%", fig.show='hold', echo=FALSE}
knitr::include_graphics(c(
  "data/Manchester/Plots/Growth_Results/growth_util_egal_components_number_comparisonManchester.png",
  "data/Manchester/Plots/Growth_Results/growth_util_egal_components_gcc_comparisonManchester.png"
))
```

# Disussion and conclusions

>>>>

---
title: Road Segment Prioritization for Bicycle Infrastructure
date: "`r Sys.Date()`"
geometry: margin=2cm
bibliography:
  - mybibfile.bib
  - rlbibfile.bib
#linenumbers: true
numbersections: true
abstract: "Understanding the motivators and deterrents to cycling is essential for creating infrastructure that gets more people to adopt cycling as a mode of transport. The aim of this paper is to develop methods to support the prioritisation of cycling infrastructure and cycling network design, accounting for cyclist preferences and the growing emphasis on 'filtered permeability' and 'Low Traffic Neighborhood' interventions internationally. Distance decay, routing, and network analysis techniques are used to examine where future cycling demand is most likely to arise, how such demand could be accommodated within existing street networks, and how to ensure a fair distribution of investment. The city of Manchester is used as a case study, although the methods can be applied to any city, as demonstrated in a reproducible code and open access data accompanying the paper. We find that flexible routing profiles that are adjustable to local needs can avoid road segments that are stressful to cycle on, enabling the identification of networks that balance directness against the need for safe and stress-free routes. We assess two different policy interventions , 'utilitarian' and 'egalitarian', by analyzing the connectivity of the city through community detection. From a policy perspective, our findings indicate that the 'egalitarian' prioritisation model approach is the more adequate to support cost-effective strategic cycle network planning. \\par\\textbf{Keywords:} cycling networks, low-traffic neighborhoods, routing, transport equity"
# keywords: "cycling, low-traffic neighborhoods, potential demand, routing"
output:
  # bookdown::github_document2:
  bookdown::pdf_document2:
  # # bookdown::word_document2:
      toc: false
      extra_dependencies: ["subcaption", "longtable", "float", "multirow", "placeins", "multicol"]
editor_options: 
  markdown: 
    wrap: 72
---

@@

<!-- - Intro: possible trim down substantially then add some literature in (a) Calculating Potential Demand [DONE]/ (b) Routing [DONE]/ (c) Road Segment Prioritization -->

<!-- - Community Detection: Where in the document should this section be? It is now before Road Segment Prioritization, but I can merge them -->

<!-- Can go into road segment section. -->

<!-- - Should we use other cities for the sake of comparison (compare communities, road types etc) -->

<!-- - How to add tables properly (they are now added using latex code which only shows up when outputting to pdf) -->

<!-- - Footnotes are sometimes split between pages. How do I keep them on the same page? -->

<!-- - What is the best way to write pseudocode (Steps for Algorithm 1 and 2) -->

@@

> "Inadequate cycling infrastructure discourages cycling and wastes
> public money. Much cycling infrastructure in this country is
> inadequate. It reflects a belief, conscious or otherwise, that hardly
> anyone cycles, that cycling is unimportant and that cycles must take
> no meaningful space from more important road users, such as motor
> vehicles and pedestrians - [@departmentgearchange2020]"

@@

A considerable amount of research has been done to understand what gets people to cycle.
Segregated cycling infrastructure^[*Segregated cycling infrastructure*
refers to road space that is allocated to cyclists only, with physical
separation to protect cyclists from other modes of transport.] has been
shown to increase cycling uptake [@aldred2019impacts; @goodman2014new;
@marques2015infrastructure], with the separation from motorized
vehicles being key [@winters2011motivators]. Revealed preference of
cyclists shows that they are willing to deviate from the most efficient
routes in order to commute on safer roads [@crane2017longitudinal].
However, such deviations are only considered if they do not considerably
increase route circuitry; behaviour studies have found that the
probability of choosing a route decreases in proportion to its length
relative to the shortest route [@broach2011bicycle; @winters2010far].
Another defining feature for cycling infrastructure is how well
connected it is. Cyclists prefer cohesive infrastructure, particularly
when cycling on arterial roads with high levels of motorized traffic
[@stinson2003commuter]. The lack of well-connected cycling
infrastructure is one of the main obstacles to increasing cycling uptake
[@caulfield2012determining]. While direct and cohesive cycling networks
have been shown to positively impact cycling rates, density^[making an
area's bicycle network denser means adding more cycling routes in the
area and thereby giving cyclists more route options] of the cycling
network is also vital [@schoner2014missing].

@@

*Optimization* techniques have been used to propose improvements to
cycling networks. @mesbah2012bilevel propose a bi-level formulation to
optimize allocation of cycling lanes to the network without exceeding a
set budget. They account for the effect of cycling lanes on car traffic,
and attempt to maximize utilization of said lanes with minimal impact on
car travel times. Safeguarding against increased car traffic may be counter-productive if the goal is to create a mode-shift, as research has shown that reducing road space for cars leads to less cars on the road, a phenomenon referred to as "traffic evaporation" [@nello2020environmental].
Car usage is not considered by @mauttone2017bicycle,
who develop an optimization framework that aims to minimize the total
user cost of cycling on the network. The aggregate flow^[*flow* is
used throughout this research to refer to the cycling demand when it is
routed onto the road network. The flow on any road segment is the
cumulative demand on it, resulting from cyclists commuting between
various OD pairs] on links^[*link*
refers to a road segment throughout this research] is obtained by using
shortest paths to route existing cycling demand onto the road network,
and the solution is a proposed set of links where cycling infrastructure
should be added in order to minimize the overall travel cost of cyclists
across the network. The cost of traversing a link is given as a function
of its length and whether or not it has cycling infrastructure. The
problem has also been solved by attempting to find the minimum cost of
improving roadway links to meet a desired level of service (LOS)
[@duthie2014optimization]. In this formulation, all origin-destination
(OD) pairs need to be connected by roads that meet the desired LOS. A
limitation of these problem formulations do not explicitly solve for
continuity, which is dealt with using either a constraint specifying
that each link with a bike lane should be connected to at least one
destination [@mesbah2012bilevel], a constraint on maximum deviation
from shortest paths [@duthie2014optimization], or a discontinuity
penalty to prioritize connected road segments [@mauttone2017bicycle].

In this paper, continuity is analysed by looking at the connectivity of
the network through the graph-theoretic concept of *connected
components*. Other academics have taken this approach; @natera2019data
study the existing cycling network in terms of its disconnected
components and introduce two different algorithms to connect these
components by their most critical links and, in doing so, measure the
size of the growth of the largest connected component as a function of
the kilometers of network added. The concept of connected components is
also at the core of the methodology proposed by @olmos2020data. After
routing the cycling demand onto the network links, they use percolation
theory to filter out the links based on the aggregate flow passing through them, varying the flow threshold for
filtering to identify the minimum flow at which the whole city is
connected by a giant component. While these approaches deal with
continuity better, they look at the network as a whole when attempting
to improve it, and in doing so fail to account for equitable
distribution of infrastructure.

@@

This research attempts to propose an egalitarian framework for cycling
network design. This is done by identifying the different sub-networks
that exist within the larger network, and ensuring that each gets a fair
share of investment. Trip patterns in a city are not uniformly
distributed geographically, and *community finding* methods have been
used to partition study areas into localized areas that experience a
disproportionate number of trips within them. @akbarzadeh2018designing
use a modularity maximization approach [@blondel2008fast] on taxi trip
data to identify 7 different communities in the city of Isfahan, Iran.
An optimization problem is then formulated to connect nodes within each
community with cycling infrastructure, with the emphasis being on
connectivity within the communities, not between them. @bao2017planning
adopt a similar methodology, first identifying communities and then
using a greedy network expansion algorithm to simultaneously add links
to each community. The link with the highest benefit-cost ratio in each
community is selected, and the network is grown by adding neighboring
links to the solution until a budget limit is met. The benefit is the
flow on the link, and each link is assigned a cost based on current road
conditions.

<!-- EDIT! Add more to limitations -->

Our work builds on these *community finding* approaches by proposing a
similar greedy network expansion algorithm for cycle network expansion
within communities. We incorporate community finding methods for study
area partitioning with weighted routing to avoid links that are
stressful to cycle on. In doing so, we propose a methodology that
accounts for motivators and deterrents to cycling. We propose three
sub-methodologies that expand on some of the limitations of previous
studies. These limitations include (a) bias inherent when basing network
design solely on existing cycling demand, (b) proposing routes that may
not correspond to studies on cyclist preference and government policies,
and (3) insufficient consideration of the ethical principles underlying
the analysis. Section \@ref(calculating-potential-cycling-demand)
focuses on calculating potential cycling demand. Section \@ref(routing)
focuses on routing the demand onto the road network while accounting for
cyclist preferences and government priorities. Section
\@ref(road-segment-prioritisation) outlines a methodology for partitioning the
study area based on a community finding algorithm and routed cycling
demand. It then introduces the network expansion algorithms, and compares an approach grounded in
'egalitarianism' to one grounded in 'utilitarianism'.

@@

*Origin-Destination (OD)* data can be obtained from a range of sources,
including GPS data, household travels surveys and Census data on work
locations. In areas where observed OD data is unavailable, modelling
techniques such as spatial interaction models can be used to estimate
travel volumes between zones [@black1995spatial; @martinez2013new;
@wilson1971family]. In this paper we use open access data from the UK
census [@ONS2011flowdata], which contains aggregate statistics on
number of commuters between administrative zones --- Middle layer Super
Output Areas (MSOA) --- by mode of travel. MSOAs have an average
population of just over 8000 [@ofn2018population].^[ See
<https://wicid.ukdataservice.ac.uk/> for open access to the OD data.]
Figures \@ref(fig:potdemhistograms) and \@ref(fig:desirefacetcycling)
illustrates the proportion of trips cycled by distance and the
geographic extent of the input OD dataset used in this paper.

@@

```{r desirefacetcycling, fig.cap="Current and Potential Cycling Demand", out.width="65%", fig.pos="H", fig.align="center", echo=FALSE}
knitr::include_graphics("data/Manchester/Plots/desire_facet_cycling.png")
```

@@

The above points are accounted for by using a weighted road network for
routing. This has previously been done by multiplying all road segments
without cycling infrastructure by a fixed impedance factor
[@mauttone2017bicycle], or by assigning different weights to the road
segments proportional to the investment cost of bringing them to an
acceptable level of stress for cycling [@duthie2014optimization].
Perceived stress for cyclists has been found to be a function of vehicular traffic
volume and speed [@sorton1994bicycle], both of which vary predictably
with road type. @gehrke2020cycling build on this information to use
routing impedance factors for road segments that are based on road type
and the existence of cycling infrastructure.

@@

<!-- **ADD TABLE - THIS IS BASIC** -->

<!-- ```{r, echo=FALSE, message = FALSE} -->

<!-- weight_profiles <- readxl::read_excel("Paper/paper_tables.xlsx", sheet = "Weighting Profile") -->

<!-- knitr::kable(weight_profiles, -->

<!--              caption = "Weighting Profiles") -->

<!-- ``` -->

\begin{longtable}{lrrr}
\caption{Weighting Profiles}
\label{table:weightprofiles}\\
\multirow{2}{*}{\textbf{OSM Road Type}} & \multicolumn{3}{c}{\textbf{Weighting Profile}}                                                 \\*
                                        & \multicolumn{1}{l}{\textit{Unweighted}} & \multicolumn{1}{l}{\textit{Weighted}} & \multicolumn{1}{l}{\textit{Weighted\_2}}  \endfirsthead
Cycleway                                & 1          & 1                                     & 1                                         \\
Path                                    & 1          & 0.9                                   & 0.9                                       \\
Residential                             & 1          & 0.9                                   & 0.9                                       \\
Service                                 & 1          & 0.9                                   & 0.9                                       \\
Tertiary                                & 1          & 0.9                                   & 0.9                                       \\
Track                                   & 1          & 0.9                                   & 0.9                                       \\
Unclassified                            & 1          & 0.9                                   & 0.9                                       \\
Secondary                               & 1          & 0.8                                   & 0.8                                       \\
Primary                                 & 1          & 0.7                                   & \textcolor{red}{0}                        \\
Trunk                                   & 1          & 0.6                                   & \textcolor{red}{0}                        \\
Motorway                                & 1          & \textcolor{red}{0}                    & \textcolor{red}{0}                       
\end{longtable}

@@

1.  **Weighted**: Relatively high impedance on Primary and Trunk roads
    (to minimize cycling on them).
2.  **Weighted_2**: Avoiding Primary and Trunk Roads completely.

@@

Banning cycling flow completely on trunk and primary roads may result in
excessively circuitous paths, as seen in Figure
\@ref(fig:boxplotcircuity). When routing using the *weighted* weighting
profile in Table \@ref(table:weightprofiles), we see that shortest paths
increase by less than 5% on average from unweighted shortest paths, with
the largest increases still below 30%. When routing on primary and trunk
roads is banned (*weighted_2* profile in Table
\@ref(table:weightprofiles)), the average increase relative to
unweighted shortest paths rises to 10%, with certain locations
experiencing more significant negative effects on accessibility. Given
that cyclists will only deviate from shortest paths by a certain amount
to access better cycling infrastructure (as explained in Section
\@ref(introduction)), allowing flow on some stretches of trunk and
primary roads is necessary to insure cycling uptake and equitable access
to cycling infrastructure. In its new vision for walking and cycling,
the Department for Transport acknowledges that minimal segregated
stretches of bicycle lanes on main roads will be necessary to avoid
circuitous cycling networks [@departmentgearchange2020].

@@

It should be reiterated that the weighting profile used for routing has
been developed for the purposes of this study. It creates a hierarchy of
road preference that is grounded in cyclist preferences and government
plans to create LTNs. A wide range of weighting profiles could be used
to represent different types of cyclists and road environments, as
described in @gehrke2020cycling and @furth2016network. Sensitivity
analysis could be done to determine an optimal weighting profile, but
given the variation in city road
networks<!-- (Figure \ref{fig:perc_person-km}) -->, this would probably
require calibration to the specific city. More accurate routing could be
carried out given the availability of road-level data. In such cases we
would add additional impedance to specific roads, giving more useful
routing results than the current methodology which considers all roads
of the same type to be equivalent.

One use-case of such granular data would be to identify roads that serve
schools. The Department of Transport notes that the number of school
children being driven to school has trebled over the past 40 years
[@departmentgearchange2020], and so having cycling infrastructure
serving schools is key to achieving the government target of getting
more children to cycle. This would not be difficult, as over 75% of
children in the UK live within a 15 minute cycle from their school
[@departmentcycleinfradesign2020]. @goodman2019scenarios show that if
dutch levels of cycling were achieved in the UK, the % of children
cycling to school could increase from 1.8% to 41%. In their typology of
cyclists, @dill2013four found that a majority of people who say they
would never cycle had never cycled to school, whereas confident cyclists
were those most likely to have cycled to school. Getting people to cycle
from a young age is therefore key to achieving societal change in
commuting habits.

# Road Segment Prioritisation

@@

As explained in Section
\@ref(ethical-underpinnings-and-proposed-approach), a major challenge
facing 'top-down' planning approaches is how to incorporate egalitarian
principles by fairly distributing investments in cycling infrastructure.
One way of quantifying this is to split up the city into smaller
geospatial areas and target equal investment in each of those areas.
This approach could also help ensure that on-the-ground surveys are made
by local stakeholders, an important component of the planning process
[@parkin2018designing]. Community detection offers us a way to
delineate such a split; cyclists are limited in their commuting distance
(see Figure \@ref(fig:cyclinghistmanc)), and so trip attractors are more
likely to have a local catchment area of cyclists.

@@

<!-- 1. Create links between MSOA centroids and weigh these links by potential cycling demand between them.
2. Use Louvian method to determine optimal number of communities and assign each MSOA centroid to a community.
3. Assign each road link to the same community as the closest MSOA centroid to it. -->

```{r, eval = FALSE}
1. Create links between MSOA centroids and weigh these links by potential cycling demand 
   between them.
2. Use the Louvain method to determine the optimal number of communities and assign each MSOA 
   centroid to a community.
3. Assign each road link to the same community as the closest MSOA centroid to it.
```

@@

```{r communitiesmanchester, fig.cap="Communities Based on Potential Cycling Demand (Manchester)", out.width="75%", fig.show='hold', fig.align = "center", echo=FALSE}
knitr::include_graphics("data/Manchester/Plots/communities_alternative_Manchester.png")
```

@@

## Algorithm 2: Egalitarian Expansion

The first algorithm focuses on connectivity and directness, but not on
fairly distributing investment. The latter is not a requirement for
increasing cycling uptake, but it is fundamental for spatial equity, as
explained in Section \@ref(ethical-underpinnings-and-proposed-approach). This algorithm incorporates
the ideal of fair distribution by using community detection to partition
the road network.

@@

The results of the community detection are used to evaluate the
algorithms. This is done by looking at the *person-km satisfied* as
cycling infrastructure is added. Person-km is a measure of the total km
cycled on a road segment, so it is the product of the number of
potential commuters cycling on that road segment ($flow$) and the length
of the segment in km ($l$). For each road segment, the person-km is
equal to $flow * l$. In the case of Manchester, Table
\@ref(tab:personkmtable) shows that almost half of the person-km is in
community 1 , while only 0.5% of total person-km on the network is in
community 5.

@@

Looking at the person-km satisfied (Figure \@ref(fig:growthtotal)), we
see that the incremental addition of cycling infrastructure is better
distributed between communities using Algorithm 2; equal distribution of
investment results in the gain in % of person km satisfied in each
community being inversely correlated with the size of the community. In
addition, we find that the restrictions imposed by Algorithm 2 on the
network expansion do not seem to have a noticeable effect on the
city-wide % of person-km satisfied. Comparing both algorithms, we can
see that Algorithm 1 provides only marginally quicker city-wide gains
than Algorithm 2.

```{r growthtotal, fig.cap="Comparing Overall (Dashed) and Community Level Person-Km Satisfied (Manchester)", out.width="35%", fig.show='hold', fig.align = "center", echo=FALSE}
knitr::include_graphics(c(
  "data/Manchester/Plots/Growth_Results/growth_utilitarian_satisfied_km_both_flow_column.png",
  "data/Manchester/Plots/Growth_Results/growth_egalitarian_satisfied_km_both_flow_column.png"
))
```

@@

```{r growth3MapandBar, fig.cap="Road Segment Priority (left), disaggregated by road type (right) - Egalitarian Growth", out.width="35%", fig.show='hold', fig.align="center", echo=FALSE}
knitr::include_graphics(c(
  "data/Manchester/Plots/Growth_Results/growth_egalitarian_priority_all_FLOW.png",
  "data/Manchester/Plots/Growth_Results/growth_egalitarian_investment_highways_flow.png"
))
```

@@

```{r componentsandGCC, fig.cap="Network Characteristics", out.width="30%", fig.show='hold', fig.pos="H", fig.align = "center", echo=FALSE}
knitr::include_graphics(c(
  "data/Manchester/Plots/Growth_Results/growth_util_egal_components_number_comparisonManchester.png",
  "data/Manchester/Plots/Growth_Results/growth_util_egal_components_gcc_comparisonManchester.png"
))
```
